Modelling the emergence of whisker barrels

### 2020-06-02 paper/briefpaper/brief.tex
Tweaks on re-reading

A cortical boundary enclosing 41 corresponding barrels was traced from a
cytochrome oxidase stain from \cite{zheng_signal_2001}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes. From random initial conditions for
$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
within \cmnt{$\approx 20000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
(see Fig.\,\ref{fig:main}D). \cmnt{In a chosen region, the ratio of the connection density
  of the TC type $i$ which makes the densest connections to the sum of all
  connection densities gives a measure of the localization of the connection
density. We computed this measure across the entire barrel domain and named it $\omega$
(Fig.\,\ref{fig:main}D). $\omega$ increases during the first $20000$ simulation
steps and converges on a stable value. Fig.\,{fig:main}E shows
the localization measure plotted as a luminance map, giving a more graded
impression of the structure which develops in the system.}

@@

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the maximally
  localized solution, which for $D$ should be small. Thus, there is a
  trade-off between good localization and good Voroni pattern conformation. We
  chose the parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A to present in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii) and iii) indicate the effect
  of varying $D$ in either direction from its optimum. ii) is a less
  exaggerated version of the pattern in i), whereas iii) shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

The most involved part of this computation is to find the divergence of the
flux of axonal branching, $\mb{J}_i(\mb{x},t)$, the term in parentheses in
Eq.\,\ref{eq:da}:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{blue}+
  \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
\end{equation}
%
\cmnt{where we have defined $\hat{a}_i=\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
\sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.  Because the divergence
operator is distributive, Eq.\,\ref{eq:divJ} can be expanded using vector
calculus identities (dropping references to $\mb{x}$ and $t$ for clarity):
%
\begin{equation}
\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
\nabla\vcdot\big(a_i \mb{g}_i\big) \color{blue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
\end{equation}
%
Applying the vector calculus product rule identity yields
%
\begin{equation} \label{eq:divJExpanded}
%
\nabla\vcdot\mb{J}_i =
%
D \dvrg a_i % term1
-
a_i\nabla\vcdot\mb{g}_i % term2
-
\mb{g}_i\vcdot\nabla a_i % term3
\color{blue}
+ \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
+ \frac{\epsilon}{N-1} \nabla \hat{a}_i\vcdot\nabla{a}_i% term1_2
\color{black}
,
\end{equation}
%
which has \cmnt{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
$a_i$); ii)~a time-independent modulator of $a_i$ (because
$\nabla\vcdot\mb{g}_i$ is a time-independent static field); iii)~the scalar
product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
\cmnt{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
  of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
means of Gauss's Theorem following \cite{lee_hexagonal_2014}.

>>>>

A cortical boundary enclosing 41 corresponding barrels was traced from a
cytochrome oxidase stain from \cite{zheng_signal_2001}, and
Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
resulting domain, using $M=2$ linear signalling gradients aligned with the
anterior-posterior and medial-lateral axes. From random initial conditions for
$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
within \cmnt{$\approx 20000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
(see Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of \emph{the connection density
  of the TC type $i$ which makes the densest connections} to \emph{the sum of all
  connection densities} gives a measure of the localization of the connection
density. We computed this measure across the entire barrel domain and named it $\omega$
(see Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first $20000$ simulation
steps and converges on a stable value. Fig.\,\ref{fig:main}E shows
the localization measure plotted as a luminance map, giving a more graded
impression of the structure which develops in the system.}

@@

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the maximally
  localized solution, which for $D$ should be small. Thus, there is a
  trade-off between good localization and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii) and iii) indicate the effect
  of varying $D$ in either direction from its optimum. ii) is a less
  exaggerated version of the pattern in i), whereas iii) shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

The most involved part of this computation is to find the divergence of the
flux of axonal branching, $\mb{J}_i(\mb{x},t)$, the term in parentheses in
Eq.\,\ref{eq:da}:
%
\begin{equation}
  \label{eq:divJ}
  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{blue}+
  \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
\end{equation}
%
\cmnt{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
\sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.  Because the divergence
operator is distributive, Eq.\,\ref{eq:divJ} can be expanded using vector
calculus identities (dropping references to $\mb{x}$ and $t$ for clarity):
%
\begin{equation}
\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
\nabla\vcdot\big(a_i \mb{g}_i\big) \color{blue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
\end{equation}
%
Applying the vector calculus product rule identity yields
%
\begin{equation} \label{eq:divJExpanded}
%
\nabla\vcdot\mb{J}_i =
%
D \dvrg a_i % term1
-
a_i\nabla\vcdot\mb{g}_i % term2
-
\mb{g}_i\vcdot\nabla a_i % term3
\color{blue}
+ \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
+ \frac{\epsilon}{N-1} \nabla \hat{a}_i\vcdot\nabla{a}_i% term1_2
\color{black}
,
\end{equation}
%
which has \cmnt{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
$a_i$); ii)~a time-independent modulator of $a_i$ (because
$\nabla\vcdot\mb{g}_i$ is a time-independent static field); iii)~the scalar
product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
\cmnt{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
  of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
means of Gauss's Theorem following \cite{lee_hexagonal_2014}.
